\chapter{Introduction}
\label{intro}

HiCEnterprise predicts long-range interactions between regions/domains based on Hi-C maps.
Establish contact! Please visit the program web page to see examples of the running the package: \url{http://regulomics.mimuw.edu.pl/wp/hicenterprise/}

\section{What is in the package?}
HiCEnterprise takes a list of small (1-3 bins) regions or Topological Associated Domains and searches for their significant interactions on given Hi-C maps.
Package consists of two parts that perform different types of analysis: prediction of interactions between regions and domains.

\subsection{Types of analyses}

\subsubsection{Regions}
This part of the package is an implementation of the method for identification of long-range interacting regions and
creating interaction profiles based on Hi-C data as introduced by (Won et al. 2016).

More in~\ref{regions}.

\subsubsection{Domains}
In the domain part of the package we implemented the method for prediction of long-range interactions between TADs
(Topologically Associating Domains) as presented in (Niskanen et al. n.d.).

More in~\ref{domains}.

\section{Dependencies}
\begin{itemize}
\item \href{https://www.python.org}{Python} version 2.7 or 3.5/6
\item required python packages: numpy, scipy, statsmodels, matplotlib.
\item if you want to plot through R, you need have R installed with needed packages.

\end{itemize}

\section{Basic usage}

\begin{lstlisting}
HiCEnterprise type options
\end{lstlisting}


'type' can be either regions or domains. Other options are type dependent.

If you want to see the list of available arguments with short explanations type either \\
\lstinline{HiCEnterprise regions --help}
or \lstinline{HiCEnterprise domains --help}. Full specifications for each argument can be found in the documentation
(\ref{regions} and~\ref{domains}) \\


\subsection{Installation}

\begin{lstlisting}
python setup.py install
\end{lstlisting}


You may need administrative privileges to do this.\\


\subsubsection{Testing the installation}

To test the scripts, please install pytest package, go to the main HiCEnterprise directory and run:

\begin{lstlisting}
python setup.py test
\end{lstlisting}


\subsection{Examples}

We provided some example files (used for testing too) with which you can learn how to use the program. In order to easy download package here are available small example files that are useful for testing if the program works properly. In order to test program on the real Hi-C maps, please download example files from the HiCEnterprise web page: \url{http://regulomics.mimuw.edu.pl/wp/hicenterprise/}
Here you can see two example runs:

\subsubsection{Regions}

\begin{lstlisting}
HiCEnterprise regions regions_options
\end{lstlisting}


\subsubsection{Domains}

\begin{lstlisting}
HiCEnterprise domains domains_options
\end{lstlisting}


Additional examples and specifications for each argument can be found in the documentation for a specific type
(\ref{regions} and~\ref{domains}).

